### 2019-07-15 paper/ms.tex
Remove unecessary section

\documentclass{aastex62}

\usepackage{graphicx}
\usepackage{gensymb}
\usepackage{bm}
\usepackage{amsmath}

\newcommand{\Gus}[1]{\textcolor{red}{#1}}

\newcommand{\Msun}{\ensuremath{\text{M}_\odot}}
\newcommand{\pc}{\ensuremath{\text{pc}}}
\newcommand{\kpc}{\ensuremath{\text{kpc}}}
\newcommand{\Mpc}{\ensuremath{\text{Mpc}}}
\newcommand{\Myr}{\ensuremath{\text{Myr}}}
\newcommand{\Gyr}{\ensuremath{\text{Gyr}}}
\newcommand{\SFR}{\ensuremath{\text{SFR}}}

\newcommand{\MHz}{\ensuremath{\text{MHz}}}
\newcommand{\GHz}{\ensuremath{\text{GHz}}}

\newcommand{\CI}{\ensuremath{\text{C~\textsc{i}}}}
\newcommand{\CII}{\ensuremath{\text{C~\textsc{ii}}}}
\newcommand{\OIII}{\ensuremath{\text{O~\textsc{iii}}}}
\newcommand{\CO}{\ensuremath{\text{CO}}}

\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\uth}{\textsuperscript{th}}
\newcommand{\n}{\text{n}}

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}

\newcommand{\ps}[1]{\ensuremath{P_{#1,#1}}}
\newcommand{\xps}[2]{\ensuremath{P_{#1,#2}}}

\newcommand{\kmax}{\ensuremath{k_{\text{max}}}}

\newcommand{\mynu}[2]{\ensuremath{\nu_{#1,\text{#2}}}}
\newcommand{\denps}{\ensuremath{P_{\delta,\delta}}}
\newcommand{\pstot}[1]{\ensuremath{P_{#1,\text{tot}}}}
\newcommand{\Bsum}{\ensuremath{\tilde{B}}}

\newcommand{\kpar}{k_{\parallel}}
\newcommand{\kperp}{\bm{k}_{\perp}}
\newcommand{\apar}{\alpha_{\parallel}}
\newcommand{\aperp}{\alpha_{\perp}}

\newcommand{\E}[1]{\mathrm{E}[#1]}
\newcommand{\Var}[1]{\mathrm{Var}[#1]}
\newcommand{\Cov}[2]{\mathrm{Cov}[#1,#2]}
\newcommand{\avg}[1]{\ensuremath{\langle #1 \rangle}}
\newcommand{\SN}{\ensuremath{\text{S}/\text{N}}}

\newcommand{\bi}{\ensuremath{\avg{b_i}}}
\newcommand{\bj}{\ensuremath{\avg{b_j}}}
\newcommand{\Ii}{\ensuremath{\avg{I_i}}}
\newcommand{\Ij}{\ensuremath{\avg{I_j}}}

\newcommand{\penn}{Department of Physics \& Astronomy, University of
Pennsylvania, 209 South 33rd Street, Philadelphia, PA 19104, USA}
\newcommand{\harvard}{Center for Astrophysics {\normalfont |} Harvard \& Smithsonian, 60
Garden Street, Cambridge, MA 02138, USA}

\begin{document}

\title{Multi-line Intensity Mapping}

\correspondingauthor{Angus Beane}
\email{abeane@sas.upenn.edu}

\author{Angus Beane}
\affil{\harvard}
\affil{\penn}

\author{Adam Lidz}
\affil{\penn}

\author{others}

\shorttitle{Multi-line Intensity Mapping}
\shortauthors{Beane et al.}

\begin{abstract}
% Context
Intensity mapping of galactic emission lines stands to shed light on both
galactic astrophysics and cosmology from $z\sim1$ to $z\sim10$ in the coming
years. However, several observational challenges must first be met, including
the problems of interloper and foreground contamination.
% Aims
Cross-correlations of two different emission lines can confront these
challenges. By combining the cross-correlations between three or more lines,
one can measure the large-scale bias encoded in the auto-spectrum of each
line. We aim to understand the general case in more detail.
% Methods
Using the Fisher matrix formalism, we develop a framework for understanding
the observational constraints possible with intensity mapping of an arbitrary
number of lines and their associated cross-spectra.
% Results
Results.
% Conclusions
We find that multi-line intensity mapping is a promising technique, and future
intensity mapping efforts should take care to maximize the shared area between
different surveys.
\end{abstract}

\section{Introduction}\label{sec:intro}

\section{Technical Background}\label{sec:tech_back}
We first lay out the approach to modelling the average intensity of an
arbitrary emission line, the spatial clustering of that signal, the
cross-spectrum between two different lines, and the fisher matrix approach for
estimating the bias factors from an arbitrary number of cross-spectra or
cross-spectra and auto-spectra.

\subsection{Intrinsic Line Emission}\label{ssec:line_emission}
The average intensity of an arbitrary line is calculated according
to the simple model of \citet{2016ApJ...825..143L}, which is based upon
\citet{2011ApJ...741...70L} and \citet{2013ApJ...768...15P}. The details are laid out in
Appendix~\ref{app:co_int}, but in brief the model assumes a \citet{1976ApJ...203..297S} form for the star formation rate function and a linear relationship
between a halo's luminosity and its star formation rate with constants of
proportionality from \citet{2010JCAP...11..016V}.

We note that in this model the redshift evolution of each line is identical up
to the luminosity-SFR constant of proportionality. In reality, we do not
expect this to be the case. For example, the different CO lines should have
varying intrinsic strengths since the typical phase of the ISM in each
emitting galaxies should change as a function of redshift. This highlights one
of the limitations of the line intensity model adopted here. Fully
understanding the underlying galactic population will require more nuanced
models, though the best path forward is not clear.

To model the intrinsic spatial fluctuations of each line, we follow
\citet{2016ApJ...825..143L} and \citet{2016ApJ...832..165C}, though we adopt
the notation of the former. We use the word ``intrinsic'' here not to mean the
signal in the cosmological rest frame, but rather the observed signal assuming
the observed coordinates have been transformed to comoving coordinates with
the correct redshift. This is not the case for interloping lines, for which
their contribution to the signal is distorted because the assumed redshift and
the emitted redshift are not the same.

The intrinsic emission from a line $i$ is given by
\beq\label{eq:int_ps}
P_{i,i}(k, \mu, z) = \avg{I_i}^2 \avg{b_i}^2 (1 + \beta_i \mu^2)^2 
              D\left[\mu k \sigma_p(z)\right] P_{\delta, \delta}(k, z)\text{,}
              % + P_{i, \text{shot}}\text{,}
\eeq
where $\mu=k_{\parallel}/k$ is the cosine of the angle between the wavevector
$\bm{k}$ and the line of sight direction, $\avg{I_i}$ is the average specific
intensity of line $i$, and $\avg{b_i}$ is the average luminosity-weighted bias
of the emitting galaxies. The Kaiser effect \citep{1987MNRAS.227....1K} is implemented
through the $(1 + \beta_i \mu^2)^2$ factor, where $\beta_i =
f_{\Omega}/\avg{b_i}$ with $f_{\Omega} = \text{d}\ln{D}/\text{d}\ln{a}$ is the
logarithmic derivative of the growth factor, well approximated by $f_{\Omega}
\approx \Omega^{0.55}$ \citep{2005PhRvD..72d3529L}. We assume a Lorentzian form for the
finger-of-god suppression:
\beq\label{eq:fog}
D(x) = \frac{1}{1+x^2}\text{.}
\eeq
We approximate the pairwise velocity dispersion by $\sigma_p(z) =
\sigma_v(z)/\sqrt{2}$ with $\sigma_v^2(z)$ being the variance of the
line-of-sight component of the velocity field according to linear theory.
Equation~\eqref{eq:int_ps} neglects the shot noise component of the power
spectrum, which is an acceptable omission for the large scales considered in
this work. Our neglect to add the shot noise component is out of convenience considering the scales we will encounter, but this is not to imply that the shot noise component is uninteresting in its own right ({\bf cite some?}) even in the cross-spectra ({\bf cite Breysse 2019}). We use the \citet{1998ApJ...496..605E} model for the linear theory
matter power spectrum, $P_{\delta,\delta}(k, z)$. All basic cosmological
calculations are performed using the code \texttt{colossus} \citep{2018ApJS..239...35D}.

Throughout this work we will only consider the angle-averaged power spectrum, that is:
\beq\label{eq:int_ps}
\begin{split}
P_{i,i}(k, z) &= \avg{I_i}^2 \avg{b_i}^2 C_i(k) P_{\delta, \delta}(k, z)\\
                   &\equiv B_i^2 C_i(k) P_{\delta, \delta}(k, z)\text{,}
              % + P_{i, \text{shot}}\text{,}
\end{split}
\eeq
where we have defined $C_i(k)$ to be the amplification of the power spectrum
coming from the average over the angle-dependent terms in
Equation~\eqref{eq:int_ps}, and $B_i$ to be the product of $\avg{b_i}$ and
$\avg{I_i}$. Henceforth all power spectra will be assumed to be angle
averaged.

For the cross-spectrum between two lines $i$ and $j$, we assume that
\beq\label{eq:cross_spec}
P_{i,j}(k, z) = r_{i,j}(k, z) \sqrt{P_{i,i}(k, z) P_{j,j}(k, z)}\text{,}
\eeq
where $r_{i,j}$ is the cross-correlation coefficient between line $i$ and $j$.
We will only consider scales where $r_{i,j} = 1$.

\subsection{Multi-line Formalism}\label{ssec:multi_line}
Given a set of cross-spectra and their covariance matrix, we can, in
principle, determine the optimal constraints that can be placed on both the
average bias $\avg{b_i}$ and average intensity $\avg{I_i}$. However, such a
procedure requires measuring the angular dependence of the power spectrum
({\bf cite}), which may be fairly challenging \citep{2019arXiv190500209C}.
Therefore, after fixing the value of $b_i$ for each line $i$, we only consider
the constraint that can be placed on each $B_i$. This is slightly
inconsistent, since in principle we should first determine the constraint on
$\bi$ and $\Ii$ and then convert this into a constraint on $B_i$. But our
approximation is reasonable because {\bf of some reason}.

We can write down the Fisher matrix as
\beq\label{eq:fisher}
F_{i,j} = 
\int \frac{\text{d}^3k}{(2\pi)^3} V_{\text{surv}}
% \int \frac{\text{d}k\,k^2}{2\pi^2} V_{\text{surv}} 
\sum_{l} \sum_{m}
\frac{\partial \hat{P}_{l}(\bm{k})}{\partial B_i}
\left(\Cov{\hat{P}_{l}(\bm{k})}{\hat{P}_{m}(\bm{k})}\right)^{-1}
\frac{\partial \hat{P}_{m}(\bm{k})}{\partial B_j}\text{,}
\eeq
where the sum over $l$ and $m$ each is over each \emph{pair} of lines. Since
we will be considering only angle-averaged power spectra, we can use the
angle-averaged form of the Fisher matrix:
\beq\label{eq:fisher}
F_{i,j} = 
\int \frac{\text{d}k\,k^2}{2\pi^2} V_{\text{surv}}
% \int \frac{\text{d}k\,k^2}{2\pi^2} V_{\text{surv}} 
\sum_{l} \sum_{m}
\frac{\partial \hat{P}_{l}(k)}{\partial B_i}
\left(\Cov{\hat{P}_{l}(k)}{\hat{P}_{m}(k)}\right)^{-1}
\frac{\partial \hat{P}_{m}(k)}{\partial B_j}\text{.}
\eeq

In
order to compute the covariance matrix, we use the result that \citep[e.g.][]{2010JCAP...11..016V},
\beq\label{eq:var_cov}
\begin{split}
\Var{\xps{i}{j}} &= \xps{i}{j}^2 + \pstot{i}\pstot{j} \\
\Cov{\xps{i}{j}}{\xps{i}{k}} &= \pstot{i}\xps{j}{k} +
\xps{i}{j}\xps{i}{k}\text{,}
\end{split}
\eeq
where $\pstot{i} = \ps{i} + N_i$. The computation of the noise $N_i$ is given in the next subsection.

\subsection{Noise Components}\label{ssec:noise}
We describe the computation of each noise component. In principle the noise
should consist of three components:
\beq\label{eq:noise_comp}
N_i = N_{i,\text{instrumental}} + N_{i,\text{interlopers}} + N_{i,\text{foreground}}\text{.}
\eeq
However, in this work we will ignore the foreground contribution. Our work
should then be seen as a best-case scenario. {\bf add more words}

\subsubsection{Instrumental Noise}\label{sssec:instrumental}
{\bf Add.}

\subsubsection{Interloping and Extraloping Lines}\label{sssec:interlopers}
The noise for interloping and extraloping lines can be written as
\citep{2016ApJ...825..143L, 2016ApJ...832..165C},
\beq\label{eq:ptot_intlpr}
N_{i,\text{interlopers}} = \sum_{j} \frac{1}{\apar(z_j) \aperp(z_j)^2} 
            P_j\left(\frac{\kpar}{\apar(z_j)}, 
                     \frac{\kperp}{\aperp(z_j)}\right)\text{,}
\eeq
where the sum on $j$ is over all interloping lines, $z_j$ is the
emitted redshift of line $j$, and where,
\begin{alignat}{2}
\apar = \frac{H(z_i)}{H(z_j)} \frac{1+z_j}{1+z_i} \quad \quad &&
\aperp = \frac{D_{A}(z_j)}{D_A(z_i)}\text{,}
\end{alignat}
where $z_i$ is the emitted redshift of line $i$ and $D_A(z)$ is the comoving
angular diameter distance to redshift $z$.

\section{Three Field Approach}\label{sec:tf}
We begin by considering the three field approach. We first start with this
relatively simple case to gain some intuition for the \SN{} behavior when the
cross-spectra of multiple lines are combined.

We are concerned with the problem of estimating the power spectrum of each
line only from the three cross-spectra. To be more concrete, we model each
auto-spectrum and cross-spectrum as,
\beq\label{eq:auto_xps}
\begin{split}
\ps{i}(k,z) &= B_i(z)^2 \denps(k,z) \\
\xps{i}{j}(k,z) &= B_i(z) B_j(z) \denps(k,z)\text{,}
\end{split}
\eeq
where $B_i(z) \equiv b_i(z) \avg{I_i}(z)$. We will generally suppress the
redshift and $k$ arguments for brevity. We can model the auto-spectra and
cross-spectra in this way only on sufficiently large scales, i.e. only when
$k<\kmax$ for some value of \kmax{}. The value of \kmax{} is set by two
conditions. First, we must have that linear biasing is a good description for
each field $i$ so that $B_i$ for each line is independent of $k$. Second, for
each pair of lines $(i,j)$ we must have that $\abs{r_{i,j}(k)} = 1$ when
$k<\kmax$.

Eventually it will be desirable to place constraints on $\denps(k)$. For now,
however, we simply wish to understand how well future experiments will be able
to determine each of $B_i$ as a function of $z$. We can proceed simply by
computing the relevant Fisher matrix:
\beq\label{eq:fisher_tf}
F_{i,j} = 
\int \frac{\text{d}^3k}{(2\pi)^3} V_{\text{surv}}
\sum_{l} \sum_{m}
\frac{\partial \hat{P}_{l}(\bm{k})}{\partial B_i}
\left(\Cov{\hat{P}_{l}(\bm{k})}{\hat{P}_{m}(\bm{k})}\right)^{-1}
\frac{\partial \hat{P}_{m}(\bm{k})}{\partial B_j}\text{,}
\eeq
where the sum over $l$ and $m$ each is over every \emph{pair} of lines.
Equation~\ref{eq:fisher_tf} can be readily modified to accomodate the case of
more than three lines (by extending the summations to include these lines) and
to accomodate the auto-spectra (by extending the sum over $l$ and $m$ to
include them). We will consider these cases in Sections~\Gus{x}~and~\Gus{y},
respectively.

For the variance and covariance of the power spectra, it can be shown that,
\beq\label{eq:var_cov}
\begin{split}
\Var{\xps{i}{j}} &= \xps{i}{j}^2 + \pstot{i}\pstot{j} \\
\Cov{\xps{i}{j}}{\xps{i}{k}} &= \pstot{i}\xps{j}{k} +
\xps{i}{j}\xps{i}{k}\text{,}
\end{split}
\eeq
where $\pstot{i} = \ps{i} + N_i$. Note that $N_i$ contains contributions
from instrument noise, residual foreground contamination, and
interloping/extraloping lines. While computing the Fisher matrix numerically
is trivial, we will first work out the noise-dominated case which has a
somewhat tractable analytic solution.

\subsection{White Noise-dominated Regime} \label{ssec:tf_noisedom}
In the white noise-dominated regime, we have that each $N_i \gg \ps{i}$ and
$N_i(k) = N_i$. The relevant variance and covariance formulae then reduce to,
\beq\label{eq:var_cov}
\begin{split}
\Var{\xps{i}{j}} &= N_iN_j \\
\Cov{\xps{i}{j}}{\xps{i}{k}} &= N_i\xps{j}{k}\text{.}
\end{split}
\eeq
Since the covariance formula is only first order in $N_i$, we will ignore
contributions from terms like it and only consider terms involving
$\Var{\xps{i}{j}}$. In such a case, the formula for the Fisher matrix
simplifies greatly and we can write it down as:
\beq \label{eq:fmat_tf_exp}
F = V_k^2
\begin{pmatrix}
\frac{B_2^2}{N_1N_2}+\frac{B_3^2}{N_1N_3} & \frac{B_1B_2}{N_1N_2} & \frac{B_1B_3}{N_1N_3} \\
\frac{B_2B_1}{N_2N_1} & \frac{B_3^2}{N_2N_3}+\frac{B_1^2}{N_2N_1} & \frac{B_2B_3}{N_2N_3} \\
\frac{B_3B_1}{N_3N_1} & \frac{B_3B_2}{N_3N_2} & \frac{B_1^2}{N_3N_1}+\frac{B_2^2}{N_3N_2}
\end{pmatrix}
\text{,}
\eeq
where,
\beq \label{eq:Vk}
V_k^2 \equiv V_{\text{surv}} \int \frac{\text{d}k\,k^2}{2\pi^2} \denps(k)^2\text{.}
\eeq

Inverting this matrix we find that the covariance matrix of the bias factors is,
\beq \label{eq:covmat_tf_wn}
C = F^{-1} = \frac{1}{4V_k^2}
\begin{pmatrix}
\frac{B_1^2N_2N_3 + B_2^2N_3N_1 + B_3^2N_1N_2}{B_2^2B_3^2} & \frac{-B_1^2N_2N_3 - B_2^2N_3N_1 + B_3^2N_1N_2}{B_1B_2B_3^2} & \frac{-B_1^2N_2N_3 + B_2^2N_3N_1 - B_3^2N_1N_2}{B_1B_2^2B_3} \\
\frac{-B_1^2N_2N_3 - B_2^2N_3N_1 + B_3^2N_1N_2}{B_1B_2B_3^2} & \frac{B_1^2N_2N_3 + B_2^2N_3N_1 + B_3^2N_1N_2}{B_3^2B_1^2} & \frac{B_1^2N_2N_3 - B_2^2N_3N_1 - B_3^2N_1N_2}{B_1^2B_2B_3} \\
\frac{B_1^2N_2N_3 - B_2^2N_3N_1 - B_3^2N_1N_2}{B_1^2B_2B_3} & \frac{B_1^2N_2N_3 - B_2^2N_3N_1 - B_3^2N_1N_2}{B_1^2B_2B_3} & \frac{B_1^2N_2N_3 + B_2^2N_3N_1 + B_3^2N_1N_2}{B_1^2B_2^2}
\end{pmatrix}
\text{.}
\eeq
Examining the diagonal components we have that:
\beq \label{eq:var_bias_tf_wn}
\begin{split}
C_{1,1} = \sigma_{B_1}^2 &= \frac{1}{4V_k^2} \left( B_1^2 \frac{N_2N_3}{B_2^2B_3^2} + N_1(\frac{N_2}{B_2^2} + \frac{N_3}{B_3^2}) \right) \\
\implies \frac{\sigma_{B_1}^2}{B_1^2} &= \frac{1}{4V_k^2} \left( \frac{N_1N_2}{B_1^2B_2^2} + \frac{N_2N_3}{B_2^2B_3^2} + \frac{N_3N_1}{B_3^2B_1^2} \right)\text{.}
\end{split}
\eeq
The symmetry of this result implies that,
\beq \label{eq:frac_error_same}
\frac{B_1}{\sigma_{B_1}} = \frac{B_2}{\sigma_{B_2}} = 
\frac{B_3}{\sigma_{B_3}} \text{.}
\eeq
Unfortunately the surprising result that the fractional error on each line is
the same only holds in the three field case.

\begin{figure}[h!]
\begin{center}
\includegraphics{fig/corner_Nconstant.pdf}
\end{center}
\caption{A toy corner plot showing the degeneracies between different lines
with different intrinsic \SN{} ratios under the white-noise dominated
assumption. We set $B_1^2 = 1$, $B_2^2 = 2$, and $B_3^2 = 3$ along with $N_1,
N_2, N_3 = 100$ and $V_k^2 = 2\times10^6$. Since each line has the same value
of $N$, the \SN{} of each line is driven by the value of $B_i$, with $B_3$
having the highest intrinsic \SN{} and $B_1$ the lowest. We plot each axis as
a fraction of the true value for each of $B_1$, $B_2$, and $B_3$, and see that
the marginalized posterior on each bias factor is the same (\textit{diagonal
panels}) as we expect from Equation~\ref{eq:frac_error_same}. There is a
strong negative covariance between $B_2$ and $B_3$, a mildly negative
covariance between $B_3$ and $B_1$, and a mildly positive covariance between
$B_1$ and $B_2$. An explanation for this behavior is given in the text.}
\label{fig:tf_Nconst}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics{fig/corner_Nconstant_Bhigher.pdf}
\end{center}
\caption{We perform the same procedure as in Figure~\ref{fig:tf_Nconst}, except
now we slightly increase the intrinsic \SN{} of $B_2$ and $B_3$ by setting
$B_2^2 = 4$ and $B_3^2 = 9$. The values for each $N_i$ and $V_k$ are the same
as in Figure~\ref{fig:tf_Nconst}. We see that even though nothing about line
$1$ has changed (i.e. no additional measurements of line $1$ have been made),
the \SN{} on $B_1$ has improved because the fractional error on each line must
be the same (Equation~\ref{eq:frac_error_same}). The degeneracies between the
different lines discussed in Figure~\ref{fig:tf_Nconst} become stronger in
this case.}
\label{fig:tf_Nconst_Bhigher}
\end{figure}

Now we wish to consider how the \SN{} on each $B_i$ from the various
cross-spectra is related to the \SN{} in the auto-spectrum case. We first
write down the \SN{} from the auto-spectrum in the white-noise dominated
regime:
\beq \label{eq:autops_sn_wn}
\SN(B_1, \text{auto-spectrum}) = \frac{B_1^2}{N_1} V_k
\eeq

Now, let us compare the \SN{} of the two approaches in various cases. This
will be simpler to write if we introduce $\tilde{B}_i \equiv B_i^2/N_i$. Then,
\beq
\frac{\SN(B_1, \text{three-field})}{\SN(B_1, \text{auto-spectrum})} =
\frac{2}{\tilde{B}_1}
\sqrt{\frac{\tilde{B}_1\tilde{B}_2\tilde{B}_3}{\tilde{B}_1 + \tilde{B}_2 + \tilde{B}_3}}
\text{.}
\eeq
There are several interesting cases to consider with this formula. For
starters, let us assume that each field is measured to the same precision,
i.e. $\tilde{B}_1 = \tilde{B}_2 = \tilde{B}_3 = \tilde{B}$. In this case, we
have that
\beq
\frac{\SN(B_1, \text{three-field})}{\SN(B_1, \text{auto-spectrum})} =
\frac{2}{\sqrt{3}} \approx 1.15
\text{.}
\eeq
So, the three-field approach gives a marginally better \SN{} than the
auto-spectrum approach. This is encouraging, because it means that a tentative
auto-spectrum measurement could be checked against the value derived from the
three-field approach. Next, let us assume that the field $1$ has a much
stronger signal strength than $2$ or $3$ --- i.e. $\tilde{B}_1 \gg
\tilde{B}_2,\tilde{B}_3$. In this case we have that
\beq
\frac{\SN(B_1, \text{three-field})}{\SN(B_1, \text{auto-spectrum})} =
2\frac{\sqrt{\tilde{B}_2\tilde{B}_3}}{\tilde{B}_1}\text{.}
\eeq
In this case we see that it is unlikely for the $\text{S}/\text{N}$ on $B_1$
to be higher in the three-field case than in the auto-spectrum case. On the
other hand, let us consider the case where one of the other lines is measured
to much higher precision, e.g. $\tilde{B}_2 \gg \tilde{B}_1,\tilde{B}_3$. In
this case, we have that
\beq
\frac{\SN(B_1, \text{three-field})}{\SN(B_1, \text{auto-spectrum})} =
2\sqrt{\frac{\tilde{B}_3}{\tilde{B}_1}}\text{.}
\eeq
As long as field $1$ and $3$ are measured to comparable precision, a gain in
the \SN{} of $\sim2$ is possible with the three-field approach. Finally, let
us consider the most interesting case in which fields $2$ and $3$ are measured
to much higher precision than field $1$ --- i.e. $\tilde{B}_2, \tilde{B}_3 \gg
\tilde{B}_1$. For simplicity let us also assume that $\tilde{B}_2 \approx
\tilde{B}_3 = \tilde{B}$. In this case, we have that
\beq
\frac{\SN(B_1, \text{three-field})}{\SN(B_1, \text{auto-spectrum})} =
\sqrt{2}\sqrt{\frac{\tilde{B}}{\tilde{B}_1}}
\text{.}
\eeq
As a result, if two lines are measured with strong precision, then we can
strongly boost the \SN{} of a third, weakly measured line with the three-field
approach.

As an illustration, we construct a corner plot in Figure~\ref{fig:tf_Nconst}
for the inferred bias values using toy values of $B_1^2 = 1$, $B_2^2 = 2$, and
$B_3^2 = 3$ along with $N_1, N_2, N_3 = 100$ and $V_k^2 = 2\times10^6$. The
units here are arbitrary. Since each line has the same value of $N$, the \SN{}
of each line is driven by the value of $B_i$, with $B_3$ having the highest
intrinsic \SN{} and $B_1$ the lowest. We make the same assumptions that go
into Equation~\ref{eq:fmat_tf_exp}, i.e. that we are white-noise dominated in
each line. By plotting each axis as a fraction of the true value for each of
$B_1$, $B_2$, and $B_3$, we see that the marginalized posterior on each bias
factor is the same (\textit{diagonal panels}). This is equivalent to saying
that the fractional error on each bias factor is the same for each line.

Moving on to the joint panels in Figure~\ref{fig:tf_Nconst}, we see that there
is a strong negative covariance between $B_2$ and $B_3$, a mildly negative
covariance between $B_3$ and $B_1$, and a mildly positive covariance between
$B_1$ and $B_2$. We can intuitively understand this behavior by considering
that each cross-spectrum places a prior on the product $B_1B_2$, $B_2B_3$, and
$B_3B_1$, with the width of that prior given by the \SN{} of that
cross-spectrum.

In this toy scenario, the product $B_2B_3$ will have the highest \SN{} and
therefore the tightest prior, and therefore to account for a larger $B_3$ one
must have a smaller $B_2$ (and vice-versa), hence the strong negative
correlation. We can make this argument for the relationship between $B_2$ and
$B_3$ since $B_2B_3$ has the tightest prior, and therefore the products
$B_1B_2$ and $B_3B_1$ are not relevant, since they have wider priors. The next
tighest prior is on the product $B_3B_1$. Increasing the value of $B_1$ means
that we must decrease the value of $B_3$ to keep $B_3B_1$ constant. Since
$B_2B_3$ has a tighter prior, this means that we must also decrease the value
of $B_2$. The prior on $B_1B_2$ is irrelevant, since it is wider than the
other two.\footnote{Although in this case, to keep $B_1B_2$ constant while
increasing $B_1$ we would need to decrease $B_2$ anyways.}

Finally, we turn to the product $B_1B_2$, which has the widest prior of the
three products. If we increase $B_1$, then naively we would need to decrease
$B_2$. However, increasing $B_1$ also means that we must decrease $B_3$ to
keep $B_3B_1$ constant. Furthermore, decreasing $B_3$ means that we must
increase $B_2$ to keep $B_2B_3$ constant. Since the priors on $B_2B_3$ and
$B_3B_1$ are tighter than on $B_1B_2$, these constraints dominate, and as a
result increasing $B_1$ means that we must increase $B_2$. Therefore, $B_1$
and $B_2$ are positively correlated.

In Figure~\ref{fig:tf_Nconst_Bhigher} we perform the same procedure, except
now we slightly increase the intrinsic \SN{} of $B_2$ and $B_3$ by setting
$B_2^2 = 4$ and $B_3^2 = 9$. Notice that we have not changed the intrinsic
\SN{} of field $1$. Two things are apparent. First, the strength of the
correlations that we discussed for Figure~\ref{fig:tf_Nconst} are stronger,
since the different priors have become more disparate in their relative
tightness. Second, as a consequence of the fractional \SN{} being the same for
each line, we see that the error on $B_1$ has become much smaller (and
similarly for $B_2$ and $B_3$). However, this is particularly interesting in
the case of $B_1$. No modification has been made to it's intrinsic \SN{} (or,
in other words, no additional measurements have been made of line $1$).
Nonetheless, the \SN{} on the inferred $B_1$ has become higher.

\section{Random Thoughts}
Like doing a matched filter??? Trying to understand intuitively why this
works. If you have a super well measured field, then its like you're God who
knows that the density field is. Its almost like using that density field as a
matched filter?

\appendix
\section{CO Intensity Modelling} \label{app:co_int}
To model the statistical fluctuation and amplitude of the \CO{} intensity maps
we follow \citet{2016ApJ...825..143L}, and briefly summarize the procedure
here. We begin by assuming a Schechter form for the star formation rate
function,
\beq\label{eq:schechter}
\phi(\SFR)\,\text{d}\SFR = \phi_*
\left(\frac{\SFR}{\SFR_*}\right)^{\alpha}\exp{\left[-\frac{\SFR}{\SFR_*}\right]}
\frac{\text{d}\SFR}{\SFR_*}
\eeq
where $\alpha$ is the faint-end slope, $\SFR_*$ is the characteristic
star-formation rate, and $\phi_*$ is the characteristic number density. The
average specific intensity in a line $l$ can be related to the comoving
emissivity by \citep{2011ApJ...741...70L, 2013ApJ...768...15P},
\beq\label{eq:emiss_to_int}
\avg{I_l} = \frac{\epsilon_l}{4\pi \nu_{\text{rest},l}}\frac{c}{H(z)}\text{,}
\eeq
where $\nu_{\text{rest},l}$ is the rest-frame emission frequency and
$\epsilon_l$ is the comoving emissivity, each of line $l$. Assuming a linear
luminosity-SFR relation with a constant of proportionality of $L_0^l$, along
with the Schechter form for the SFR function (Equation~\ref{eq:schechter}), it
follows that \citep{2013ApJ...768...15P},
\beq\label{eq:com_emiss}
\epsilon_l = \phi_* L_0^{l}
\frac{\SFR_*}{1\,\Msun\text{yr}^{-1}}\Gamma{(2+\alpha)}\text{.}
\eeq
We adopt the value of $L_0^{\CO}$ for each line transition from
\citet{2010JCAP...11..016V}.

\bibliography{references}

\end{document}